\propertitle

### 2019-11-19 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T08:53:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ is the size of the pool of all neurons that could just
as well have been recorded by our measuring procedure. \enquote{Just as
  well} indicates equal degrees of belief. Whether the equality of such
beliefs is justified, or for which sampling procedures it could be
justified, is the fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe; and similar. $N$ can only be assessed case by
case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
might be enough for qualitative inferences. As discussed in the next
section, even if the exact number $N$ were known, the maximum-entropy
distribution ought to be interpreted qualitatively or semi-quantitatively.

If our uncertainty about $N$ spans several orders of magnitude, we can of
course assign a probability to the possible values of $N$ based on our
background information and the frequency data $\yf$. The procedure is
similar to that of \sect~\ref{sec:marginalization} and appendix***. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
$N$ is the exponential of the relative entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) \approx \exp\bigl\{
  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto p(N)\;\exp\bigl\{
  T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is a simple illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming five
moments $M$ are sufficient,
formula~\eqref{eq:prob_data_from_N} gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Our uncertainty is about a scale factor, so it can be represented by equal
pre-data probabilities $1/5$ about these partial orders of magnitude. Thus
from~\eqref{eq:data1_f_from_N} we find 
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T14:47:34+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is formula~\eqref{eq:equal_m_factorialmoments}, on which
our method is based, and which asserts the equality of the factorial
moments of the full network and those of the sample. This formula is only
valid if our beliefs about the activity of the sample $\pxx(\ya)$ and that
of the full network $\pxx(\yA)$ are related by a hypergeometric
distribution:
\begin{equation}
  \label{eq:hypergeometric_relation}
    \pxx(\ya)=
  \binom{n}{\ya}\;\binom{N-n}{\yA-\ya}\;\binom{N}{\yA}^{-1}
\;\pxxx(\yA),
\end{equation}
characteristic of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, $N$ is the size of the pool of all neurons that could just
as well have been recorded by our measuring procedure. \enquote{Just as
  well} indicates equal degrees of belief. Whether the equality of such
beliefs is justified, or for which sampling procedures it could be
justified, is the fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe; and similar. $N$ can only be assessed case by
case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
magnitude might be enough for qualitative inferences. As discussed in the
next section, even if the exact number $N$ were known, the maximum-entropy
distribution ought to be interpreted qualitatively or semi-quantitatively.

If our uncertainty about $N$ spans several orders of magnitude, we can of
course assign a probability to the possible values of $N$ based on our
background information and the frequency data $\yf$. The procedure is
similar to that of \sect~\ref{sec:marginalization} and appendix***. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
$N$ is the exponential of minus the relative entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto
  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is a simple illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
of first five moments is sufficient, formula~\eqref{eq:prob_data_from_N}
gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Since our uncertainty regards a scale factor, it can be represented by
equal pre-data probabilities $1/5$ about these partial orders of magnitude.
Thus from~\eqref{eq:data1_f_from_N} we find
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

In this way we can make inferences -- for example about the sufficiency of
a set of moments, or about the marginal sample distribution -- that take
into account our uncertainty about $N$. We must make sure to avoid
circularities, though: for example, we can't assume a set of moments $M$ to
be sufficient and assess our uncertainty about $N$ conditional on it, and
then use this uncertainty to assess the sufficiency of $M$. But we can
approximately assess the sufficiency of a subset of moments $M' \subset M$
taking into account the uncertainty of $N$ conditional on $M$.

A rigorous assessment would involve a more expensive, full Bayesian
calculation; but such calculation would make the whole maximum-entropy
approach superfluous. We discuss this final point in the next section.
